Applied Temporal Analysis: A Complete Run of the FraCaS Test Suite

### 2021-04-22 paper2020/paper.tex
all JP todos addressed

Thanks to the unicity of action of $write(smith,x,...)$ (the subject
and direct object are fixed) we find \([t_1,t_2] = [t_3,t_4]\), and
due to the years 1991 and 1992 being disjoint we obtain contradiction.

However, the testsuite instructs that we should \emph{not} be able to refute
problem (280), with the justification that ``wrote a novel'' is a repeatable
accomplishment:

@@

We treat verbs in the progressive form as different semantically from
the non-progressive form. For example, ``John was writing a book'' is
encoded as
$∃(t_1, t_2). t_1≤ t_2, t_2 ≤ now, PROG\_write(John,book,t_1,t_2)$,
while ``John wrote a book'' is encoded as
$∃(t_1, t_2). t_1≤ t_2, t_2 ≤ now, write(John,book,t_1,t_2)$. This distinction is
necessary because in our analysis the progressive form ($PROG\_write$) is subject to
$\constant{subsumption}$. That is, if John is writing in the interval
$[t_1,t_2]$ then he is writing in any sub-interval of
$[t_1,t_2]$. This interpretation corresponds to the idea the the action takes place continuously over the whole interval.
However, the same cannot be said of the non-continuous form ($write$): the
end-points of the interval indicate the time needed to complete the
achievement. (For example, ``John wrote a book in 1993'' neither
entails ``John wrote a book in January 1993'' nor ``John wrote a book
in December 1993''.) (In fact, \emph{write}, in the non-progressive
from, is on the contrary subject to $\constant{unicity}$.) Finally, we
also have $write(x,y,t_1,t_2) → PROG\_write(x,y,t_1,t_2)$. That is,
the achievement (or \emph{activity} in our terminology) variant implies the stative variant, for the
same interval.  Consequently we get the entailment from ``John wrote a
book in 1993'' to ``John was writing a book in 1993'', but not the other way around.

@@

Of particular note is the use of the pronoun ``it'', and the fact that
adverbials expressions such that ``in 1992'' are lexicalized.  We also
follow the GF convention to postfix lexical items with the name of
their category. Most of the other categories follow usual naming
conventions. We remind the reader that ``slash'' categories are used
to swap the order of arguments (compared to non-slashed categories of
similar names).

>>>>

Thanks to the unicity of action of $write(smith,x,...)$ (the subject
and direct object are fixed) we find \([t_1,t_2] = [t_3,t_4]\), and
due to the years 1991 and 1992 being disjoint we obtain contradiction. In sum, no special notion of accomplishment is needs to be invoked: we only need the principle of unicity of action.

Yet, the testsuite instructs that we should \emph{not} be able to refute
problem (280), with the justification that ``wrote a novel'' is a repeatable
accomplishment:

@@

We treat verbs in the progressive form as different semantically from
the non-progressive form. For example, ``John was writing a book'' is
encoded as
$∃(t_1, t_2). t_1≤ t_2, t_2 ≤ now, PROG\_write(John,book,t_1,t_2)$,
while ``John wrote a book'' is encoded as
$∃(t_1, t_2). t_1≤ t_2, t_2 ≤ now, write(John,book,t_1,t_2)$. This distinction is
necessary because in our analysis the progressive form ($PROG\_write$) is subject to
$\constant{subsumption}$. That is, if John is writing in the interval
$[t_1,t_2]$ then he is writing in any sub-interval of
$[t_1,t_2]$. This interpretation corresponds to the idea that the action takes place continuously over the whole interval.
However, the same cannot be said of the non-continuous form ($write$): the
end-points of the interval indicate the time needed to complete the
achievement. (For example, ``John wrote a book in 1993'' neither
entails ``John wrote a book in January 1993'' nor ``John wrote a book
in December 1993''.) (In fact, \emph{write}, in the non-progressive
from, is on the contrary subject to $\constant{unicity}$.) Finally, we
also have $write(x,y,t_1,t_2) → PROG\_write(x,y,t_1,t_2)$. That is,
the achievement (or \emph{activity} in our terminology) variant implies the stative variant, for the
same interval.  Consequently we get the entailment from ``John wrote a
book in 1993'' to ``John was writing a book in 1993'', but not the other way around.

@@

Of particular note is the use of the pronoun ``it'', and the fact that
adverbial expressions such that ``in 1992'' are lexicalized.  We also
follow the GF convention to postfix lexical items with the name of
their category. Most of the other categories follow usual naming
conventions. We remind the reader that ``slash'' categories are used
to swap the order of arguments (compared to non-slashed categories of
similar names).
